---
title: 'filtering'
author: 'Mick McQuaid'
date: '2020-02-01'
output:
    pdf_document:
    latex_engine: xelatex
header-includes:
  \usepackage{booktabs}
---

```{r echo=FALSE}
# This chunk is just to make it possible to shrink the typeface in succeeding chunks. Mainly this will be used for the crosstabs.
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

# Initialization

```{r initialize}
library(tidyverse)
library(data.table)
nyc311<-fread("mini311.csv")
names(nyc311)<-names(nyc311) %>%
  stringr::str_replace_all("\\s", ".")
```

# Creating the main data frame

The following chunk creates everything we need for a stacked bar chart. There
is one problem, though. If we try to sort the bars by length, we'll fail
because the complaints are per borough. We need a column with total complaints
for all boroughs.

```{r filtering}
df1 <- nyc311 %>%
	group_by(Complaint.Type,Borough) %>%
	subset(select=c(Complaint.Type,Borough)) %>%
	summarize(Complaints = n()) %>%
	filter(Complaints > 100)
```

# Joining data frames

The quickest way (not perhaps the most elegant) to get a column of total
complaints for all boroughs is to group by complaint type, add the complaints
per complaint type, then join the result to the data frame created above.
Warning: only do `full_join()` with very small data frames or you will run out of memory
before the join is completed.

```{r joining}
df2 <- df1 %>%
	group_by(Complaint.Type) %>%
	summarize(totComplaints=sum(Complaints))

df3<-full_join(df1,df2)
```

# Creating the barchart

Now we can create the barchart and order the bars by the `totComplaints`
column, where there is an identical entry for each `Complaint.Type`/`Borough`
pair. After you run this code, look at `df1`, `df2`, and `df3` if you're at all
confused by what is going on.

```{r barchart}
df3 %>%
	ggplot( aes(x=reorder(Complaint.Type,totComplaints), y=Complaints, fill=Borough)) +
    xlab("Category") +
    geom_bar(stat="identity") +
    coord_flip() +
    ggtitle("Top Complaints by Type and Borough")


