---
title: "Basics"
author: "Mick McQuaid"
date: "3/10/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load the tidyverse packages.

```{r step one}
library(tidyverse)
```

Display the built-in data set, mpg.

```{r step two}
mpg
```

Show a scatter plot of displ times hwy (which is the same as saying engine size times highway gas mileage).

```{r step three}
ggplot(data=mpg)+
  geom_point(mapping = aes(x=displ,y=hwy))
```

Show the exact same scatter plot.

```{r step four}
ggplot(data=mpg,mapping = aes(x=displ,y=hwy))+
  geom_point()
```

Show the exact same scatter plot.

```{r step five}
ggplot(mpg,aes(x=displ,y=hwy))+
  geom_point()
```

Show the exact same scatter plot.

```{r step fiveb}
ggplot(mpg,aes(x=displ,y=cty))+
  geom_point(mpg,mapping=aes(x=displ,y=hwy))
```

Show a slightly different scatter plot.

```{r step fivec}
ggplot(mpg,aes(x=displ,y=hwy))+
  geom_point(mpg,mapping=aes(x=displ,y=cty))
```

Show the hwy scatter plot except in color.

```{r step six}
ggplot(mpg,aes(x=displ,y=hwy))+
  geom_point(mapping=aes(color=class))
```

Show the exact same scatter plot.

```{r step seven}
ggplot(mpg,aes(x=displ,y=hwy))+
  geom_point(aes(color=class))
```

Show the scatter plot in the book where the two-seaters are highlighted in red.

```{r step eight}
ggplot(mpg,aes(x=displ,y=hwy))+
  geom_point()+
  geom_point(data=mpg[mpg$class=="2seater",],color="red")
```

In the above code, I selected rows of the data set. Specifically, I selected all rows where class is "2seater". I colored just those rows red.
