---
title: 'concise plotting tips'
author: 'Mick McQuaid'
date: '2019-03-11'
output:
    pdf_document:
header-includes:
  \usepackage{booktabs}
---

```{r echo=FALSE}
knitr::opts_chunk$set(echo=TRUE)
```


```{r echo=FALSE}
# This chunk is just to make it possible to shrink the typeface in succeeding chunks. Mainly this will be used for the crosstabs.
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

# Initialization

```{r initialize}
library(tidyverse)
library(data.table)
library(xtable)
```

# File Reading

```{r readmini}
samplemini<-fread("mini311.csv")
```

# Example from Maitha

```{r barplot}
maxAgency <- samplemini %>%
  group_by(Agency) %>%
  summarize(count=n()) %>%
  filter(count>0)
  maxAgency$Agency<-factor(maxAgency$Agency,
  levels=maxAgency$Agency[order(maxAgency$count)])

ggplot(data = maxAgency, mapping = aes(x=Agency,y=count))+
  geom_bar(stat="identity", fill = "#87CEFA") +
  ggtitle("Complaints per Agency")+
  xlab("Agency")+
  ylab("Number of Complaints")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.background = element_rect(fill = "transparent",colour = "gray"),
        plot.title = element_text(hjust = 0.5))+
  coord_flip() +
  labs(caption = "(based on data from NYC 311)")
```

# Example with diagonal labels

```{r diagonal x labels}
ggplot(data = maxAgency, mapping = aes(x=Agency,y=count))+
  geom_bar(stat="identity", fill = "#87CEFA") +
  ggtitle("Complaints per Agency")+
  xlab("Agency")+
  ylab("Number of Complaints")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.background = element_rect(fill = "transparent",colour = "gray"),
        plot.title = element_text(hjust = 0.5))+
  theme( axis.text.x = element_text(angle = 45, hjust = 1))+
  labs(caption = "(based on data from NYC 311)")
```

# Example using slow ggplotting

This example uses a technique called *slow ggplotting*, popularized by a guide called *The ggplot flipbook*, available at
[https://evamaerey.github.io/ggplot_flipbook/ggplot_flipbook_xaringan.html](https://evamaerey.github.io/ggplot_flipbook/ggplot_flipbook_xaringan.html)
.
This technique adds each element one at a time. You can comment out all but the first element and run the chunk, then add one element at a time, rerunning the chunk to see the effect of adding each element separately.

```{r slow ggplotting}
ggplot(maxAgency)+
  aes(x=Agency)+
  aes(y=count)+
  geom_bar(stat="identity", fill = "#87CEFA") +
  labs(title="Complaints per Agency")+
  labs(subtitle="New York City agencies handling 311 complaints")+
  labs(x="Agency")+
  labs(y="Number of Complaints")+
  labs(caption = "(based on data from NYC 311)")+
  theme(panel.grid.major = element_blank())+
  theme(panel.grid.minor = element_blank())+
  theme(plot.background = element_rect(fill = "transparent",colour = "gray"))+
  theme(plot.title = element_text(hjust = 0.5))+
  theme( axis.text.x = element_text(angle = 45, hjust = 1))
```
