---
# This is a sample homework iii starter file. You need to do more than this for full credit.
# At least you should fill in all the places with parentheses for a B and use your imagination to add more for an A.
# render this file by saying
#   library(rmarkdown)
#   render("ii.Rmd",pdf_document(latex_engine="xelatex"))
# at the R console
title: 'homework iii'
author: 'Ayush Kumar Shah'
date: '2020-09-22'
output:
    pdf_document:
    latex_engine: xelatex
header-includes:
  \usepackage{booktabs}
---

```{r echo=FALSE}
# This chunk is just to make it possible to shrink the typeface in succeeding chunks. Mainly this will be used for the crosstabs.
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```


# Introduction
In this report, we try to explore the nyc311 data and find answers about the 
data. Some of the questions we try to answer in this exploration are:

- What are the most frequent categories of complaints?
- How is the frequent complaint categories distributed across different Boroughs?
- How does the overall status of the complaints vary across different categories?
- What are the status of the most frequent complaint categories?
- Which agencies are the top 10 largest responding City Government Agencies?

We perform various transformations to the data using the `dplyr` package to
answer these queries. We also use `ggplot` to visualize the results and also the
distributions of different variables of the data.

# Initialization
Here we load the tidyverse packages and the `data.table` package and load the nyc311 data set. Then we fix the column names of the nyc311 data so that they have no spaces.

```{r initialize}
library(tidyverse)
library(data.table)
nyc311<-fread("311_Service_Requests_from_2010_to_Present.csv")
names(nyc311)<-names(nyc311) %>%
stringr::str_replace_all("\\s", ".")
mini311<-nyc311[sample(nrow(nyc311),10000),]
write.csv(mini311,"mini311.csv")
```

# Working with maps
## Reading the saved shorter sample of the data 
This is done since original data has too many points to visualize.

```{r readmini}
sample<-fread("mini311.csv")
```

## Selecting a single complaint type "Noise"
```{r selection}
complaintlocs <- sample %>%
  select(Complaint.Type,
    Longitude,
    Latitude
  )
noisecompl <- complaintlocs %>%
  filter(Complaint.Type == "Noise")
```

## Displaying the map
```{r generatemap}
library(ggmap)
key <- "api_key"
register_google(key=key)
nyc_map <- get_map(location=c(lon=-73.9,lat=40.75),
		   maptype="terrain",zoom=10)
map <- ggmap(nyc_map) +
  geom_point(data=noisecompl,aes(x=Longitude,y=Latitude),
	     size=0.4,alpha=0.2,color="red") +
  ggtitle("Map of Sample") +
  theme(plot.title=element_text(hjust=0.5)) +
  xlab("Longitude") + ylab("Latitude")
map
```

# Most frequent Complaint Categories

Let's view the Top 10 most frequent categories of the complaints registered 
along with the count and count %.

```{r top10_complaints}
top10_complaints <- 
  nyc311 %>%
  group_by(Complaint.Type) %>%
  summarize(count = n()) %>%
  mutate(rank = min_rank(desc(count)),
         'proportion in %' = count / sum(count) * 100) %>%
  filter(rank <= 10) %>%
  arrange(rank) %>%
  select(-rank)

top10_complaints
```

## Pie chart of the most frequent complaint categiories

Let's plot the counts generated above in a pie chart.

```{r top10_complaints_plot}
top10_complaints_plot <- 
  ggplot(top10_complaints, aes(x="", y=count, fill=Complaint.Type)) +
  geom_bar(stat = "identity") +
  coord_polar("y", start=0)

top10_complaints_plot
```

# Most freqeuent Complaint Categories Count across different Boroughs

Now, let's view the counts of the top 10 frequent complain categories 
across different Boroughs using a facet plot.

```{r complaint_boroughs}
complaint_types <- nyc311 %>%
  group_by(Complaint.Type, Borough) %>%
  summarize(Complaints = n()) %>%
  filter(Complaint.Type %in% top10_complaints$Complaint.Type,
         Borough != 'Unspecified')

ct_borough <- ggplot(complaint_types) + 
  geom_bar(stat="identity", 
           aes(x=Complaint.Type, y=Complaints, fill=Complaint.Type),
           show.legend = FALSE) + 
  facet_wrap(~ Borough) +
  coord_flip() +
  xlab("Complaint Type") +
  ggtitle("Top 10 Complaints Count by Category across different Boroughs")

ct_borough
```
## Visualizing using geom_tile

```{r geom_tile1}
nyc311 %>%
  filter(Complaint.Type %in% top10_complaints$Complaint.Type,
          Borough != 'Unspecified') %>%
  count(Complaint.Type, Borough) %>%
  ggplot(mapping = aes(x = Borough, y = Complaint.Type)) +
  geom_tile(mapping = aes(fill = n))

```

# Status of complaints

## Overall status of the complaints

```{r status}
nyc311 %>% 
  group_by(Status)%>%
  summarize(count = n()) %>%
  arrange(desc(count))
```

```{r status_plot}
ggplot(data = nyc311, aes(x = Status, fill=Status)) +
  geom_bar(show.legend = FALSE) +
  coord_cartesian(ylim = c(0, 1)) +
  coord_flip() 
```
We can see that most of the status categories have very few count. So, we only
consider the 3 major categories for analysis further.

## Status of the top 10 frequent complaint categories

We are only interested in the 3 major status of the most frequent complaint
categories.

```{r status_complaints}
status_complaints <- 
  nyc311 %>% 
  group_by(Complaint.Type, Status) %>%
  summarize(Complaints = n()) %>%
  filter(Complaint.Type %in% top10_complaints$Complaint.Type,
         Status %in% c('Open', 'Closed', 'Pending'))

## Plotting the status of complaints
status_complaints_plot <- ggplot(status_complaints) + 
  geom_bar(stat="identity", 
           aes(x=Complaint.Type, y=Complaints, fill=Complaint.Type),
           show.legend = FALSE) + 
  facet_wrap(~ Status) +
  coord_flip() +
  xlab("Complaint Type") +
  ggtitle("Status of Top 10 Complaints Count by Category")

status_complaints_plot
```

## Visualizing using geom_tile

```{r geom_tile2}
nyc311 %>%
  filter(Complaint.Type %in% top10_complaints$Complaint.Type,
         Status %in% c('Open', 'Closed', 'Pending')) %>%
  count(Complaint.Type, Status) %>%
  ggplot(mapping = aes(x = Status, y = Complaint.Type)) +
  geom_tile(mapping = aes(fill = n))

```
Hence, we can see that among the top 10 complaints category, most of them are 
closed with a very few pending cases.

## Visualizing using box_plot

```{r box_plot}
nyc311 %>%
  filter(Complaint.Type %in% top10_complaints$Complaint.Type,
         Status %in% c('Open', 'Closed', 'Pending')) %>%
  ggplot(mapping = aes(x = Status, y = Complaint.Type)) +
  geom_boxplot()

```

# Top 10 Largest Responding City Government Agencies

We find out the top 10 city government agencies in terms of the largest
Service Requests (SR) with the count and proportion count (in percentage).

```{r largest_agencies}
bigAgency <- nyc311 %>%
  group_by(Agency) %>%
  summarize(complaints_count=n()) %>%
  mutate(rank = min_rank(desc(complaints_count)),
         'proportion in %' = complaints_count / sum(complaints_count) * 100) %>%
  filter(rank <= 10) %>%
  arrange(rank) %>%
  select(-rank)

bigAgency

```

Then we visualize it using a bar chart.

```{r plot_largest_agencies}
bigAgency$Agency<-factor(bigAgency$Agency,
  levels=bigAgency$Agency[order(bigAgency$complaints_count)])

p<-ggplot(bigAgency,aes(x=Agency,y=complaints_count, fill=Agency)) +
   geom_bar(stat="identity", show.legend = FALSE) +
   coord_flip()
p
```

# Conclusion

Hence, we applied exploratory data analysis on the `nyc311` data using various
data transformation and visualization techniques. We were able to answer several
questions about the data like:

- There are only 3 important status of complaints to consider in the data.
- There is a wide variance of the top complaint categories across different 
Boroughs.
- Among the top 10 complaints category, most of them are closed with a very few 
pending cases.

