---
title: "Chapter7"
author: "Ayush Kumar Shah"
date: "9/23/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Chapter 7 Tibbles

Tibbles are one of the unifying features of the tidyverse

## Coercing a dataframe to a tibble:
```{r}
library(tidyverse)
class(iris)
as_tibble(iris)
```

## Creating tibble
```{r}
tibble(
x = 1:5,
y = 1,
z = x ^ 2 + y
)
```

## Characters not supported by R also can be column names in tibble

```{r}
tb <- tibble(
`:)` = "smile",
` ` = "space",
`2000` = "number"
)
tb
```

## Creating tiblle Row-wise

```{r}
tribble(
~x, ~y, ~z,
#--|--|----
"a", 2, 3.6,
"b", 1, 8.5
)
```

## Tibble vs dataframe

### Printing
- only the first 10 rows, and all the columns that fit on screen
- each column reports its type

To print more,
```{r}
nycflights13::flights %>%
print(n = 10, width = Inf)
```
- `options(tibble.print_max = n, tibble.print_min = m)`:
  if more than m rows, print only n rows. Use
- `options(dplyr.print_min = Inf)` to always show all rows.
- Use `options(tibble.width = Inf)` to always print all columns,
  regardless of the width of the screen.

```{r}
?tibble
```


### Subsetting
```{r}
df <- tibble(
x = runif(5),
y = rnorm(5)
)
options(digits=2)
```

```{r}
df$x
```

```{r}
df[["x"]]
```

```{r}
df[[1]]
```

Using pipe

```{r}
df %>% .$x
df %>% .[["x"]]
```

To use functions not supported for tibbles: convert to dataframe
```{r}
class(as.data.frame(tb))
```


