---
title: "Chapter8"
author: "Ayush Kumar Shah"
date: "9/23/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Chapter 8 Data import with readr

- readr is imported with tidyverse

- functions in readr
  - read_csv()
  - read_tsv()
  - read_delim() (any delimiter)
  - read_fwf() (fixed width file)

```{r}
library(tidyverse)
```

## Rad_csv

read_csv different from read.csv of R.

```{r}
# heights <- read_csv("data/heights.csv")
#> Parsed with column specification:
#> cols(
#> earn = col_double(),
#> height = col_double(),
#> sex = col_character(),
#> ed = col_integer(),
#> age = col_integer(),
#> race = col_character()
#> )
```

## Inline tables
```{r}
read_csv("a,b,c
1,2,3
4,5,6")
```

## Documenting the data

```{r}
read_csv("The first line of metadata
The second line of metadata
x,y,z
1,2,3", skip = 2)
```


```{r}
read_csv("# A comment I want to skip
x,y,z
1,2,3", comment = "#")
```

Give default names to columns

```{r}
read_csv("1,2,3\n4,5,6", col_names = FALSE)
```

Pass col names
```{r}
read_csv("1,2,3\n4,5,6", col_names = c("x", "y", "z"))
```

Missing data
```{r}
read_csv("a,b,c\n1,2,.", na = ".")
```

## Advantages compared to R's read.csv

- Faster (10x), progress bar.
- produce tibbles.
- more reproducible across all type of systems.

