---
title: "Chapter9"
author: "Ayush Kumar Shah"
date: "9/23/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tidying data

- tidyr package, member of tidyverse package.

## Tidy data

1. Each variable must have its own column.
2. Each observation must have its own row.
3. Each value must have its own cell.

untidy != messy data

Most data is untidy.

Two ways of becoming untidy:

- One variable might be spread across multiple columns. Solution - tidyr: gather()
- One observation might be scattered across multiple rows. Solution - tidyr: spread()

## Gathering

```{r}
library(tidyverse)
table4a
```
Some of the column names are not names of variables, but values of a variable. 
The column names 1999 and 2000 represent values of the year variable,
and each row represents two observations, not one.

```{r}
tidy4a <- table4a %>%
gather(`1999`, `2000`, key = "year", value = "cases")
tidy4a
```

Same with table4b

```{r}
table4b
```
Some of the column names are not names of variables, but values of a variable. 
The column names 1999 and 2000 represent values of the year variable,
and each row represents two observations, not one.

### Parameters:

- The set of columns that represent values, not variables. In this
example, those are the columns 1999 and 2000.
- The name of the variable whose values form the column names.
I call that the key, and here it is year.
- The name of the variable whose values are spread over the cells.
I call that value, and here it’s the number of cases.

```{r}
tidy4b <- table4b %>%
gather(`1999`, `2000`, key = "year", value = "population")
tidy4b
```

### Left join (by dplyr)

```{r join}
left_join(tidy4a, tidy4b)
```

## Spreading

```{r}
table2
```

When an observation is scattered across multiple rows. For example, take table2—an
observation is a country in a year, but each observation is spread
across two rows:

### Parameters

- The column that contains variable names, the key column.
Here, it’s type.
- The column that contains values forms multiple variables, the
value column. Here, it’s count.

```{r spread}
spread(table2, key = type, value = count)
```

## Separating

```{r}
table3
```

one column (rate) that contains two variables (cases and population).

```{r separate}
table3 %>%
separate(rate, into = c("cases", "population"), convert=TRUE)

table3 %>%
separate(rate, into = c("cases", "population"), sep = "/")

table3 %>%
separate(year, into = c("century", "year"), sep = 2, convert=TRUE)
```


## Unite

a single variable is spread across multiple columns.

```{r unite}
table5

table5 %>%
unite(new, century, year)

table5 %>%
unite(new, century, year, sep = "")
```

## Missing values

- Explicitly (flagged with NA) - presence of an absence
- Implicitly (not present in the data) - absence of a presence


```{r}
stocks <- tibble(
year = c(2015, 2015, 2015, 2015, 2016, 2016, 2016),
qtr = c( 1, 2, 3, 4, 2, 3, 4),
return = c(1.88, 0.59, 0.35, NA, 0.92, 0.17, 2.66)
)
stocks
```

There are two missing values in this dataset:

- The return for the fourth quarter of 2015 is explicitly missing,
because the cell where its value should be instead contains NA.
- The return for the first quarter of 2016 is implicitly missing,
because it simply does not appear in the dataset.

### Making implicit missing values explicit

```{r}
stocks %>%
spread(year, return)
```

#### Using complete

```{r complete}
stocks %>%
complete(year, qtr)
```

### Making explicit missing values implicit

```{r}
stocks %>%
spread(year, return) %>%
gather(year, return, `2015`:`2016`, na.rm = TRUE)

stocks %>%
  filter(!is.na(return))
```

### Fill

```{r}
treatment <- tribble(
~ person, ~ treatment, ~response,
"Derrick Whitmore", 1, 7,
NA, 2, 10,
NA, 3, 9,
"Katherine Burke", 1, 4
)
treatment
```

Fill by most recent non missing value.

```{r fill}
treatment %>%
fill(person)
```

## Always document how you made the tidy data from the untidy data.


