---
title: "Chapter 11"
author: "Ayush Kumar Shah"
date: "10/2/2020"
output:
  html_document:
    df_print: paged
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Strings and regular expressions with stringr

```{r}
library(tidyverse)
library(stringr)
```

# Defining strings

```{r}
(string1 <- "This is a string")
(string2 <- 'To put a "quote" inside a string, use single quotes')
(string3 <- "To put a \"quote\" inside a string, use single quotes")
(string4 <- "To put a \'quote\' inside a string, use single quotes")
```

# To see the raw string:

```{r}
writeLines(string1)
writeLines(string2)
writeLines(string3)
writeLines(string4)
```

## Special characters
```{r}
?'"'
# OR
?"'"
```

# Functions

- basic: 39 string functions beginning with str_ 
- advanced: about 200 functions starting with stri_

## String Length

```{r}
str_length(c("a", "R for data science", NA))
```

## Combining Strings

```{r}
str_c("x", "y", "z")
#> [1] "xyz"
str_c("x", "y", sep = ", ")

x <- c("abc", NA)
str_c("|-", x, "-|")

str_c("|-", c("abc", "def"), "-|")

str_c("|-", str_replace_na(x), "-|")
```

## Vectors of strings

```{r}
c("x", "y", "z")
str_c(c("x", "y", "z"))
str_c(c("x", "y", "z"), collapse="")
str_c(c("x", "y", "z"), collapse=",")
str_c(c("x", "y", "z"), c("a", "b", "c"))
str_c(c("x", "y", "z"), c("a", "b", "c"), collapse="")

```

## If inside str_c
```{r}
name <- "Hadley"
time_of_day <- "morning"
birthday <- FALSE
str_c(
"Good ", time_of_day, " ", name,
if (birthday) " and HAPPY BIRTHDAY",
"."
)
```

## Substrings

```{r}
x <- c("Apple", "Banana", "Pear")
str_sub(x, 1, 3)
str_sub(x, -3, -1)
str_sub(x, 1, 1) <- str_to_lower(str_sub(x, 1, 1))
x
```

## Locales

```{r}
str_to_upper(c("i", "ı"))
#> [1] "I" "I"
str_to_upper(c("i", "ı"), locale = "tr")
```

## Regular expressions

`str_view` requires htmlwidgets and its dependencies.

```{r}
# install.packages("htmlwidgets", dependencies = TRUE)
library(htmlwidgets)
```

```{r}
x <- c("apple", "banana", "pear")
str_view(x, "an")
str_view(x, ".a.")
```
. matches any character expect enw line.
**Note: ana doesn't match since (no overlapping)**

```{r}
str_view(c("abc", "a.c", "bef"), "a\\.c")
```

```{r}
(x <- "a\\b")
writeLines(x)
str_view(x, "\\\\")
```

- ^ to match the start of the string.
- • $ to match the end of the string.

```{r}
x <- c("apple", "banana", "pear")
str_view(x, "^a")
str_view(x, "a$")
x <- c("apple pie", "apple", "apple cake")
str_view(x, "^apple$")
```


### Chaacter class and alternatives

- \d matches any digit.
- \s matches any whitespace (e.g., space, tab, newline).
- [abc] matches a, b, or c.
- [^abc] matches anything except a, b, or c.
- a | b means a or b, where a and b are strings.

```{r}
str_view(c("grey", "gray"), "gr(e|a)y")
```
### Repetitions
- ? - 0 or 1
- + - 1 or more
- * - 0 or more
- {n} - exactly n
- {n, } - n or more
- {, m} - nat most m
- {n, m} - between n and m

```{r}
x <- "1888 is the longest year in Roman numerals: MDCCCLXXXVIII"
str_view(x, "CC?")
str_view(x, "CC+")
str_view(x, 'C[LX]+')
str_view(x, "C{2}")
str_view(x, "C{2,}")
str_view(x, "C{2,3}")
```

By default these matches are “greedy”: they will match the longest
string possible. You can make them “lazy,” matching the shortest
string possible, by putting a ? after them. This is an advanced feature
of regular expressions, but it’s useful to know that it exists:

```{r}
str_view(x, "C{2,3}?")
str_view(x, 'C[LX]+?')
```

### Groups

Example: all fruits that have a repeated pair of letters:

```{r}
fruit
str_view(fruit, "(..)\\1", match = TRUE)
```

## Use of regular expressions

- Determine which strings match a pattern.
- Find the positions of matches.
- Extract the content of matches.
- Replace matches with new values.
- Split a string based on a match.

### Detect matches

```{r}
x <- c("apple", "banana", "pear")
str_detect(x, "e")
```

Number of words starting with t
```{r}
sum(str_detect(words, "^t"))
```

What proportion of common words end with a vowel?
```{r}
mean(str_detect(words, "[aeiou]$"))
```


Words without vowels : 2 ways

```{r}
# Find all words containing at least one vowel, and negate
no_vowels_1 <- !str_detect(words, "[aeiou]")
# Find all words consisting only of consonants (non-vowels)
no_vowels_2 <- str_detect(words, "^[^aeiou]+$")
identical(no_vowels_1, no_vowels_2)

str_subset(words, "^[^aeiou]+$")

# this fails
# !str_subset(words, "[aeiou]")
```

### Subset using RE (str_subset)

Equivalent to str_detect plus indexing

```{r}
words[str_detect(words, "x$")]
#> [1] "box" "sex" "six" "tax"
str_subset(words, "x$")
#> [1] "box" "sex" "six" "tax"
```

### In tibbles

```{r}
df <- tibble(
word = words,
i = seq_along(word)
)
df %>%
filter(str_detect(words, "x$"))
```

`seq_along` - position of the word in the list.

### Count

**Note: Every operation involves non overlapping matches**

```{r}
x <- c("apple", "banana", "pear")
str_count(x, "a")
str_count(x, "ana")
```

Number of vowels per word:

```{r}
mean(str_count(words, "[aeiou]"))
```

With mutate

```{r}
df %>%
mutate(
vowels = str_count(word, "[aeiou]"),
consonants = str_count(word, "[^aeiou]")
)
```


### Reminder: Matches never overlap

```{r}
str_count("abababa", "aba")
str_view("abababa", "aba")
str_view_all("abababa", "aba")
```


### Extract Matches

Sentences (comes with `stringr` package)

```{r}
length(sentences)
head(sentences)
```

Colors

```{r}
colors <- c(
"red", "orange", "yellow", "green", "blue", "purple"
)
color_match <- str_c(colors, collapse = "|")
color_match
```

Since red may be in another word like ordered, add boundary in pattern

```{r}
(color_words <- str_c("\\b" , colors, "\\b", collapse="|"))
```

Color matches

```{r}
has_color <- str_subset(sentences, color_words)
matches <- str_extract(has_color, color_words)
head(matches)
```


Sentences wtih 2 colour words.
```{r}
more <- sentences[str_count(sentences, color_words) > 1]
str_view_all(more, color_match)
```


```{r}
str_extract(more, color_match)

str_extract_all(more, color_match)
```


### str_match()

Use spacy in Python instead for NLP

```{r}
noun <- "(a|the) ([^ ]+)"
has_noun <- sentences %>%
str_subset(noun) %>%
head(10)
has_noun %>%
str_extract(noun)
```

```{r}
has_noun %>%
str_match(noun)
```


Using tibbles

```{r}
tibble(sentence = sentences) %>%
tidyr::extract(
sentence, c("article", "noun"), "(a|the) ([^ ]+)",
remove = FALSE
)
```


### Replace (str_replace)

```{r}
x <- c("apple", "pear", "banana")
str_replace(x, "[aeiou]", "-")
#> [1] "-pple" "p-ar" "b-nana"

str_replace_all(x, "[aeiou]", "-")

x <- c("1 house", "2 cars", "3 people")
str_replace_all(x, c("1" = "one", "2" = "two", "3" = "three"))
```


Using backreferences to insert components of the match

Swapping 2nd and 3rd words in each sentence.

```{r}

sentences %>%
str_replace("([^ ]+) ([^ ]+) ([^ ]+)", "\\1 \\3 \\2") %>%
head(5)
```


### Splittin (str_split)

```{r}
sentences %>%
head(5) %>%
str_split(" ")
```

To matrix

```{r}
sentences %>%
head(5) %>%
str_split(" ", simplify = TRUE)
```

Split n times
```{r}
fields <- c("Name: Hadley", "Country: NZ", "Age: 35")
fields %>% str_split(": ", n = 2, simplify = TRUE)
```