---
title: "Chapter10"
author: "Ayush Kumar Shah"
date: "10/2/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Relational Data

```{r}
library(tidyverse)
library(nycflights13)
```

```{r}
?airports
?planes
?flights
?weather
?airlines
```

# 3 purposes of verbs

- mutating joins: add new variables to one data frame from matching observations in another.
- filtering joins: filter observations from one data frame
  based on whether or not they match an observation in the other
  table.
- set operations: which treat observations as if they were set
  elements.

# Primary key and Foreign Key
## Checking if the key uniquely identifies observations:

```{r}
planes %>%
  count(tailnum) %>%
  filter(n > 1)

weather %>%
count(year, month, day, hour, origin) %>%
filter(n > 1)
```

# 1. Mutating joins

- inner join `r inner_join(x, y, by="key")` : by is optional
- left outer join: `r left_join(x, y, by="key")` : by is optional
- right outer join `r right_join(x, y, by="key")` : by is optional
- full outer join `r full_join(x, y, by="key")` : by is optional

By:
- Default: by = NULL (i.e. all columns with matching names in both tables)
- by = "somename"
- by = c("a" = "b")

# 2. Filtering joins

- `r semi_join(x, y)` - Keep all observations in x that have a match in y
- `r anti_join(x,y)` - Drop all observations in x that have a match in y

Affect rows rather than columns.

# 3. Set operations

- `r intersect(x, y)` - only observations in both x and y
- `r union(x, y)` - unique observations in x any y
- `r setdiff(x, y)` - observations in x but not in y

Treat the rows (observations) like elements in set.

DIfference from mutating joins: whole row is considered.

