---
title: "homework v"
author: "Ayush Kumar Shah"
date: "2020-10-02"
toc: true
urlcolor: blue
output:
    pdf_document:
      latex_engine: xelatex
      includes:
        in_header: header.tex
header-includes:
  \usepackage{booktabs}
---

```{r echo=FALSE}
# This chunk is just to make it possible to shrink the typeface in succeeding chunks. Mainly this will be used for the crosstabs.
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

```{r initialize, include=FALSE}
library(tidyverse)
library(data.table)
if (!require(kableExtra)){
  install.packages('kableExtra', dependencies = TRUE)
  library(kableExtra)
}
if (!require(pander)){
  install.packages('pander', dependencies = TRUE)
  library(pander)
}
```

\newpage
# Introduction

In this report, the two data sets introduced in the previous report are 
connected to the 311 
data set, using `dplyr`. Before connecting them, several operations are performed on the datasets. 

Also, few tables consisting of an extract of the data of 
each dataset as well as the final joined dataset are shown. Finally, a data
dictionary for all the data in each dataset including the final joined dataset
is also displayed. The connections between the columns will be shown in the 
next final report.

# Reading the tidied nyc311 data

The nyc311 data was tidied and saved in the previous report. In this report, the
saved tidied version of the nyc311 data set `tidied_nyc311.csv` is loaded.

A sample of 10,000 observations is also saved and used for initial computations
but replaced later by the complete dataset.

```{r read_nyc311_data}
nyc311_tidy <-fread("tidied_nyc311.csv",
              na.strings=c("","NA"))
# mini311<-nyc311[sample(nrow(nyc311),10000),]
# write_csv(mini311,"tidied_mini311.csv")
sample_tidy <- fread("tidied_mini311.csv", na.strings=c("","NA"))
```

# Loading the additional datasets

```{r read_other_data, echo = TRUE, results = 'hide', message=FALSE}
nyc_popn <- read_csv('Projected_Population_2010-2040_-_Total_By_Age_Groups.csv')

## 2005-15 graduation
# nyc_grad <- read_csv('https://data.cityofnewyork.us/resource/qk7d-gecv.csv') 
nyc_grad <- read_csv('2005_-_2011_Graduation_Outcomes_-_Borough_-_Ethnicity.csv')
names(nyc_grad)<-names(nyc_grad) %>%
stringr::str_replace_all("\\s", ".")
```

The two additional datasets introduced in the previous report are:

1. Projected Population 2010-2040 - Total By Age Groups: [Source](https://data.cityofnewyork.us/City-Government/Projected-Population-2010-2040-Total-By-Age-Groups/97pn-acdf)
2. 2005 - 2011 Graduation Outcomes - Borough - Ethnicity: [Source](https://data.cityofnewyork.us/Education/2005-2011-Graduation-Outcomes-Borough-Ethnicity/x2hp-8ukt)

## 1. [Projected Population 2010-2040 - Total By Age Groups](https://data.cityofnewyork.us/City-Government/Projected-Population-2010-2040-Total-By-Age-Groups/97pn-acdf)

Projected total New York City population for five intervals from 2010 through 
2040 by Borough, broken down by 18 age cohorts. 
(Age groups may not add up to the total due to rounding.)

This dataset is introduced so that the population information in 
the Borough can be known in correlation to the complaints in the nyc311 data.

## 2. [2005 - 2011 Graduation Outcomes - Borough - Ethnicity](https://data.cityofnewyork.us/Education/2005-2011-Graduation-Outcomes-Borough-Ethnicity/x2hp-8ukt)

Graduation results for all students by year; cohorts of 2001 through 2007 
(Classes of 2005 through 2011). Graduation Outcomes as Calculated by the New 
York State Education Department. The New York State calculation method was first
adopted for the Cohort of 2001 (Class of 2005).

Graduates are defined as those students earning either a Local or Regents 
diploma and exclude those earning either a special education (IEP) diploma or GED.

This dataset is introduced so that the educational status of the people in 
different Borough is available for further analysis in correlation with the
nyc311 data.

Both of these datasets were obtained from the 
[NYC OpenData](https://opendata.cityofnewyork.us/). 
Also, they both 
contain the column `Borough` which makes them connectable to the `nyc311` data.

# Connecting the datasets to nyc311

## Converting the additional datasets to suitable form.
Before connecting the additional datasets to the nyc311, they are converted to
suitable form.

### Gathering the years in `nycpopn`

```{r nycpopn_view, echo=FALSE}
pander(head(nyc_popn), 
       caption = "Projected_Population 2010-2040 By AgeGroups")
```

The years need to be gathered together in the `nyc_popn` data.

```{r gathering}
nyc_popn_tidy <- 
  nyc_popn %>% 
  gather('2010':'2040', key="Year", value="Population", convert = TRUE)

pander(head(nyc_popn_tidy, 10), caption = "Tidied Population data")
```

### Filtering by total population in each Borough

Since the `nyc311` has no information of Age group, so only the observations
with total population of each Borough is filtered.

```{r total_popn}
nyc_popn_tidy <- nyc_popn_tidy %>%
  filter(Age == "Total") %>%
  select(-Age)

pander(head(nyc_popn_tidy))
```

## Converting graduation dataset to suitable form

```{r nycgrad_view, echo=FALSE}
nyc_grad %>%
  select(1:6) %>%
  head() %>%
  pander(caption="2005-2011 Graduation Outcomes - Borough",
         split.table = "Inf")
```

The `Cohort.Ctaegory` is parsed as number and added with the `Cohort.Year` to
calculate the `Graduation.Year`. Then, only the relevant columns are selected.
Finally, duplicates are removed by grouping the data according to the columns,
`Graduation.Year` and `Borough`.


```{r nyc_grad_tidy}
nyc_grad_tidy <- nyc_grad %>%
  mutate(Duration = parse_number(Cohort.Category),
         Graduation.Year = Cohort.Year + Duration) %>%
  select(Graduation.Year, Borough, c(5:7)) %>%
  group_by(Graduation.Year, Borough) %>%
  summarize_all(max)
pander(head(nyc_grad_tidy), split.table = "Inf")
```
## Check Borough's values

Since the tables will be connected by `Borough`, it is checked if there are any
unspecified values in the column or mismatch in all the tables.

```{r Borough_check, echo = FALSE, warning=FALSE}
nyc311_tidy %>%
  group_by(Borough) %>%
  summarise(n = n()) %>%
  pander()

nyc_grad_tidy %>%
  group_by(Borough) %>%
  summarise(n = n()) %>%
  pander()

nyc_popn_tidy %>%
  group_by(Borough) %>%
  summarise(n = n()) %>%
  pander()
```

Since there are no Unspecified or null values, we are good to go. However, the 
values of Borough in the 2 tables table must be converted to uppercase.

## Converting to uppercase

```{r upper_case}
nyc_grad_tidy$Borough <- nyc_grad_tidy$Borough %>% 
                            str_to_upper()
nyc_popn_tidy$Borough <- nyc_popn_tidy$Borough %>% 
                            str_to_upper()
```

## Viewing the year distribution of the tables

Since these tables contain different year values, we need to sample it based on 
a particular year. So, the year distributions of all three datasets are checked.

```{r year_dist, echo=FALSE}
nyc311_tidy %>%
  ggplot() +
  geom_bar(aes(x=Created.Year)) +
  ggtitle("Created Year Distribution of Nyc311 dataset")

nyc_popn_tidy %>%
  ggplot() +
  geom_bar(aes(x=Year)) +
  ggtitle("Year Distribution of Population dataset")

nyc_grad_tidy %>%
  ggplot() +
  geom_bar(aes(x=Graduation.Year)) +
  ggtitle("Year Distribution of Graduation dataset")
```
  
The results show that `2010` would be a good year to perform the analysis since
it is common in all the tables.

## Joining the datasets

Now that all the datasets are processed, they are ready to be connected by the columns `Year` and `Borough`. Note that the column name for `Year` differs in the three tables.

Also, 2010 is selected as the year for the analysis, hence all the tables are filtered accordingly.

```{r combine}
nyc_combined <- nyc311_tidy %>%
  left_join(nyc_popn_tidy, by = c("Created.Year" = "Year", "Borough" = "Borough")) %>%
  left_join(nyc_grad_tidy, by = c("Created.Year" = "Graduation.Year", "Borough" = "Borough")) %>%
  filter(Created.Year == 2010)
```

# View extracts of the data

## Before joining

```{r view_extracts1, echo=FALSE}
pander(head(nyc311_tidy), caption = "Tidy Nyc311 data", split.cells = 5)
pander(head(nyc_popn_tidy), caption = "Projected Population 2010-2040 data (Tidy)")
pander(head(nyc_grad_tidy), caption = "2005-2011 Graduation Outcomes - Borough (Tidy)", split.table=Inf)
```

## After joining
```{r view_extracts2, echo=FALSE}
pander(head(nyc_combined), caption = "Final combined dataset", split.cells = 5)
```

\newpage
# Data Dictionary

```{r dict, echo = FALSE}
nyc_popn_info <- tribble( 
  ~ Column.Name,  ~ Description,  ~ DataType,
  "`Borough`"   , "Name of the New York City Borough" 
  , "Text",  
  "`Age`"       , "One of 18 Age cohorts like '0-4', '15-19', 'Total', and so on", "Text",
  "`Year`"  , "Year in which the population is projected", "Number",
  "`Population`"       , "The projected population value" , "Number"
)
pander(nyc_popn_info, caption="Data dictionary for Projected Population 2010-2040 dataset")

nyc_grad_info <- tribble( 
  ~ Column.Name,  ~ Description,  ~ DataType,
  "`Borough`"   , "Name of the New York City Borough", "Text",  
  "`Graduation.Year`", "The cohort's year of graduation", "Number",
  "`Total.Cohort.Num`", "Number of students in the cohort", "Number",
  "`Total.Grads.Num`", "Number of students who graduated in the cohort", "Number",
  "`Total.Grads.Pct.of.cohort`", "Percentage of students who graduated in the cohort", "Number"
)
pander(nyc_grad_info, caption="Data dictionary for 2005-2011 Graduation Outcomes dataset")

nyc_combined_info <- tribble( 
  ~ Column.Name,  ~ Description,  ~ DataType,
  "Created.Date" , "Date Service Request (SR) was created", "Floating Timestamp",
  "Created.Month" , "Month SR was created (1-12)", "Number",
  "Created.Day" , "Day of month SR was created (1-31)", "Number",
  "Created.Year" , "Year SR was created", "Number",
  "Created.Time" , "Time SR was created", "Floating Timestamp",
  "Created.Hour" , "Hour SR was created (0-24)", "Number",
  "Closed Date"  , "Date SR was closed by responding agency", "Floating Timestamp",
  "Population", "Total population of the Borough", "Number",
  "Total.Cohort.Num", "Number of students in the cohort", "Number",
  "Total.Grads.Num", "Number of students who graduated in the cohort", "Number",
  "Total.Grads.Pct.of.cohort", "Percentage of students who graduated in the cohort", "Number",
  "Agency"       , "Acronym of responding City Government Agency" , "Text",
  "Agency Name"  , "Full Agency name of responding City Government Agency", "Text",
  "Complaint Type","This is the first level of a hierarchy identifying the topic of 
                      the incident or condition.Complaint Type may have a corresponding
                      Descriptor (below) or may stand alone." , "Text",
  "Descriptor"   ,"This is associated to the Complaint Type, and provides further 
                     detail on the incident or condition.Descriptor values are dependent
                     on the Complaint Type, and are not always required in SR." , "Text",
  "Location.Type","Describes the type of location used in the address information", "Text",
  "Incident.Zip" ,"Incident location zip code, provided by geo validation.", "Text",
  "Incident.Address","House number of incident address provided by submitter.", "Text",
  "Cross.Street.1", "First Cross street based on the geo validated incident location", "Text",
  "Cross.Street.2", "Second Cross Street based on the geo validated incident location" , "Text",
  "Address.Type","Type of incident location information available." , "Text",
  "City"        , "City of the incident location provided by geovalidation.", "Text",
  "Status"      , "Status of SR submitted", "Text",
  "Due.Date"    , "Date when responding agency is expected to update the SR. 
                     This is based on the Complaint Type and internal Service Level 
                     Agreements (SLAs).", "Floating Timestamp",
  "Resolution.Description", "Describes the last action taken on the SR by the 
                               responding agency. May describe next or future steps.","Text",
  "Borough", "Provided by the submitter and confirmed by geovalidation." ,  "Text",
  "Park.Facility.Name", "If the incident location is a Parks Dept facility, the Name
                          of the facility will appear here" ,  "Text",
  "Latitude" , "Geo based Lat of the incident location" ,  "Number",
  "Longitude", "	Geo based Long of the incident location" ,  "Number"
)

landscape(knitr::kable(nyc_combined_info, caption = "Data dictionary for final Nyc combined dataset") %>%
  kable_styling(font_size=7) %>%
  column_spec(2,width="6in")
)
```

# Conclusion

Hence, the two additional datasets were joined with the `nyc311` data by converting 
the datasets into appropriate form (tidying, grouping and removal of redundant columns),
performing a left join by columns `Year` and `Borough` and finally filtering it 
by the `Year==2010` constraint.

The data extracts of tables before and after joining were were also displayed.
Finally, the data dictionary for of the two additional datasets and the final 
combined dataset was displayed.

