---
title: "homework ii"
author: "Ayush Kumar Shah"
date: '2020-09-11'
output:
  pdf_document:
    latex_engine: xelatex
header-includes: \usepackage{booktabs}
---

```{r echo=FALSE}
# This chunk is just to make it possible to shrink the typeface in succeeding chunks. Mainly this will be used for the crosstabs.
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```


# Introduction
(Introductory text should go here.)

# Initialization
Here we load the tidyverse packages and the `data.table` package and load the nyc311 data set. Then we fix the column names of the nyc311 data so that they have no spaces.

```{r initialize}
library(tidyverse)
library(data.table)
# fast for when you are starting out:
nyc311<-fread("311_Service_Requests_from_2010_to_Present.csv",nrow=10000)
# after you get going:
# nyc311<-fread("311_Service_Requests_from_2010_to_Present.csv")
# names(nyc311)<-names(nyc311) %>%
  stringr::str_replace_all("\\s", ".")
```

# Description
Here we describe the data, showing both a sample and a data dictionary.

## The head of the table
Here we produce a table of just some relevant columns of data.

```{r tabulate, results="asis"}
library(xtable)
options(xtable.comment=FALSE)
options(xtable.booktabs=TRUE)
narrow<-nyc311 %>%
  select(Agency,
	 Complaint.Type,
	 Descriptor,
	 Incident.Zip,
	 Status,
	 Borough)
xtable(head(narrow))
```

## Data Dictionary
(Here is where you s s should put the data dictionary.)

# Exploration
Here we explore the columns in the data set.

(A description of the following plot should go here.)
```{r explore}
bigAgency <- narrow %>%
  group_by(Agency) %>%
  summarize(count=n()) %>%
  filter(count>1000)
bigAgency$Agency<-factor(bigAgency$Agency,
  levels=bigAgency$Agency[order(bigAgency$count)])
p<-ggplot(bigAgency,aes(x=Agency,y=count)) +
   geom_bar(stat="identity") +
   coord_flip()
p
```

(More plots should follow here.)

Next we include a crosstabulation.

```{r crosstabs, size='footnotesize'}
xtabA<-dplyr::filter(narrow,
  Complaint.Type=='HEATING' |
  Complaint.Type=='GENERAL CONSTRUCTION' |
  Complaint.Type=='PLUMBING'
)
xtabB<-select(xtabA,Borough,"Complaint.Type")
library(gmodels)
CrossTable(xtabB$Borough,xtabB$'Complaint.Type')
```

(Some discussion of the above crosstab should follow.)

(More crosstabs or corrplots should follow.)

# Conclusion
(Tell what you did in this document here.)
